---
title: "An Explorotory Analysis of XXX"
author: "names of team members"
date: "March 4, 2018"
output: word_document
---


## Introduction

Write a brief introduction to the dataset: what it is about, how it was collected, and where you downloaded it. For example:

The following analysis uses data collected by ...  It contains... This data was downloaded from ... 


Write about the general purpose of this analysis, and list the questions you intend to address. For example:

This analysis is designed to show .... More specifically, the data analysis will answer the following questions:

1.  XXX
2.  XXX
3.  XXX

...



## 1. Data Import and Cleaning

Describe how you imported the data to tidyverse, what problems you found with the data, what steps you took to clean and tidy the data. Include R code.

#### Data Importing
Data was imported using the built in csv reading function.

```{r}
library(tidyverse)

impacts <- read.csv("impacts.csv")
orbits <- read.csv("orbits.csv")

impacts
orbits

```

One of the first steps we took to make working with the data easier was to change the column names to be shorter and easier to type.

```{r}

impacts <- rename(
  impacts,
  
  name = Object.Name,
  start = Period.Start,
  end = Period.End,
  impact = Possible.Impacts,
  cum_impact = Cumulative.Impact.Probability,
  velocity = Asteroid.Velocity,
  magnitude = Asteroid.Magnitude,
  diameter = Asteroid.Diameter..km.,
  cum_pal = Cumulative.Palermo.Scale,
  max_pal = Maximum.Palermo.Scale,
  max_tor = Maximum.Torino.Scale
  )

```

```{r}

orbits <- orbits%>%
  rename(
    name = Object.Name,
    class = Object.Classification,
    epoch = Epoch..TDB.,
    axis = Orbit.Axis..AU.,
    eccen = Orbit.Eccentricity,
    inc = Orbit.Inclination..deg.,
    ph_arg = Perihelion.Argument..deg.,
    long = Node.Longitude..deg.,
    mean = Mean.Anomoly..deg.,
    ph_dist = Perihelion.Distance..AU.,
    ah_dist = Aphelion.Distance..AU.,
    dist = Orbital.Period..yr.,
    min_int_dist = Minimum.Orbit.Intersection.Distance..AU.,
    ref = Orbital.Reference,
    mag = Asteroid.Magnitude
  )

impacts
orbits
```

The two data sets had slightly varying names. The name column for the orbit data set sometimes contained a insignificant number, with the true name in parenthesis. To clean this, we split the data set into two sets, for whether or not it has this useless number in the name.

The name also contains the year of observation in it. We split this into a seperate column using separate, and then merge orbits and impacts into one asteroids dataset, and drop the null rows.

```{r}
orbits_no_num <- filter(orbits, !grepl("\\(", name)) %>% separate(name, c("year", "name"))
orbits_no_num
orbits_num <- filter(orbits, grepl("\\(", name)) %>% separate(name, c("number", "name"), "\\(") %>% mutate(name = gsub("\\)", "", name)) %>% select(name:mag) %>% separate(name, c("year", "name"))

orbits_named <- rbind(orbits_no_num, orbits_num)
orbits_named
impacts_mod <- impacts %>% separate(name, c("year", "name"))

asteroids <- full_join(impacts_mod, orbits_named)
asteroids <- asteroids %>% drop_na()
```

## 2. Data Analysis 

May describe data tranformation you carried out before any of your analysis, and include R code.


### Question 1: 

Describe data transformation specific for this question, and include R code;

Describe how you address this specific question, and include R code;

Write your answer to the quetsion based on the analysis; may look at and cite addional sources to increase your understanding about the question.


**Note: If the question contains multiple parts (small questions), you may lable them as 1a, 1b,etc. as follows:**


#### 1a: XXX

Describe data transformation specific for this question, and include R code;

Describe how you address this specific question, and include R code;

Write your answer to the quetsion based on the analysis; may look at and cite addional sources to increase your understanding about the question.


#### 1b: XXX




### Question 2: 




### Question 3: 


...



## 3. Summary

Reflect on the exploratory analysis: 

(i)	summarize key facts learned from this analysis and discuss what you find most interesting;
(ii) may briefly discuss limitations of the analysis;
(iii)	may briefly discuss what additional data and analysis could be useful;
(iv) may discuss any other thought you have.


## References

List sources or articles you cited, if there are any.


