---
title: "An Explorotory Analysis of XXX"
author: "names of team members"
date: "March 4, 2018"
output: word_document
---


## Introduction

Write a brief introduction to the dataset: what it is about, how it was collected, and where you downloaded it. For example:

The following analysis uses data collected by ...  It contains... This data was downloaded from ... 


Write about the general purpose of this analysis, and list the questions you intend to address. For example:

This analysis is designed to show .... More specifically, the data analysis will answer the following questions:

1.  XXX
2.  XXX
3.  XXX

...



## 1. Data Import and Cleaning

Describe how you imported the data to tidyverse, what problems you found with the data, what steps you took to clean and tidy the data. Include R code.

Give a summary report of the clean dataset: how many rows and columns, variables and their data types, the first few rows of the tipple.

## 2. Data Analysis 

May describe data tranformation you carried out before any of your analysis, and include R code.


### Question 1: 

Describe data transformation specific for this question, and include R code;

Describe how you address this specific question, and include R code;

Write your answer to the quetsion based on the analysis; may look at and cite addional sources to increase your understanding about the question.


**Note: If the question contains multiple parts (small questions), you may lable them as 1a, 1b,etc. as follows:**


#### 1a: XXX

```{r}
library(tidyverse)

impacts <- read.csv("impacts.csv")
orbits <- read.csv("orbits.csv")

impacts
orbits

```

Describe data transformation specific for this question, and include R code;

Describe how you address this specific question, and include R code;

Write your answer to the quetsion based on the analysis; may look at and cite addional sources to increase your understanding about the question.


#### 1b: XXX

... 
```{r}

impacts <- rename(
  impacts,
  
  name = Object.Name,
  start = Period.Start,
  end = Period.End,
  impact = Possible.Impacts,
  cum_impact = Cumulative.Impact.Probability,
  velocity = Asteroid.Velocity,
  magnitude = Asteroid.Magnitude,
  diameter = Asteroid.Diameter..km.,
  cum_pal = Cumulative.Palermo.Scale,
  max_pal = Maximum.Palermo.Scale,
  max_tor = Maximum.Torino.Scale
  )

```

```{r}

orbits <- orbits%>%
  rename(
    name = Object.Name,
    class = Object.Classification,
    epoch = Epoch..TDB.,
    axis = Orbit.Axis..AU.,
    eccen = Orbit.Eccentricity,
    inc = Orbit.Inclination..deg.,
    ph_arg = Perihelion.Argument..deg.,
    long = Node.Longitude..deg.,
    mean = Mean.Anomoly..deg.,
    ph_dist = Perihelion.Distance..AU.,
    ah_dist = Aphelion.Distance..AU.,
    dist = Orbital.Period..yr.,
    min_int_dist = Minimum.Orbit.Intersection.Distance..AU.,
    ref = Orbital.Reference,
    mag = Asteroid.Magnitude
  )

```


### Question 2: 




### Question 3: 


...



## 3. Summary

Reflect on the exploratory analysis: 

(i)	summarize key facts learned from this analysis and discuss what you find most interesting;
(ii) may briefly discuss limitations of the analysis;
(iii)	may briefly discuss what additional data and analysis could be useful;
(iv) may discuss any other thought you have.


## References

List sources or articles you cited, if there are any.


